---
title: "Data Cleaning and Analysis"
author: "Hannah Snell and Vivienne Maxwell"
date: "2/9/2021"
output: 
  html_document:
    highlight: tango
    theme: cosmo
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: true
    df_print: kable
    
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 5, fig.height = 4, fig.align = 'center')

library(tidyverse)
library(gridExtra)
library(gdata)
library(pander)
library(mosaic)
library(kableExtra)

# import Qualtrics data 
mp2data <- read_csv("mp2data.csv")
```

# Data Cleaning and Analysis

```{r include=FALSE}
# data cleaning 
mp2data <- mp2data[-c(1, 2), ]
mp2data <- mp2data %>% 
  select(Finished, ResponseId, Q2, Q3, Q4, Q6, type, sex, climber) %>% 
  filter( 
    Finished != FALSE, 
    Q2 != "No",
    Q3 != "No, I Do Not Consent" 
    ) %>% 
  select(ResponseId, Q4, Q6, type, sex) %>% 
  mutate(type = tolower(type)) %>% 
  mutate(Q6 = as.numeric(Q6))

mp2data <- rename.vars(mp2data, from = c("Q4", "Q6", "type", "sex"),
         to = c("Climber", "Rating", "Type", "Sex"))
```

```{r echo=FALSE}
# starter plot
ggplot(mp2data, aes(x = Type, y = Rating, color = Sex)) +
  geom_point() + 
  geom_jitter(width = 0.2, height = 0.1) +
  facet_wrap(~ Climber) +
  labs(title = "Side-by-Side Dotplot: Can You Hang?", y  = "Overall Competency Rating of Climber") + 
  theme_bw()
```

```{r echo=FALSE}
# starter plot
ggplot(mp2data, aes(x = Type, y = as.numeric(Rating), color = Sex)) +
  geom_boxplot() + 
  facet_wrap(~ Climber) +
  labs(title = "Side-by-Side Boxplot: Can You Hang?", y  = "Overall Competency Rating of Climber") + 
  theme_bw()
```

```{r}
# anova data set
simple_data <- mp2data[-c(1)]
write_csv(simple_data, "simple_data.csv")
```

**ANOVA Hypothesis** 

$H_0:$ Men are perceived as physically stronger, regardless of the type of climbing. Therefore, the interaction between type of climbing (either bouldering or top roping) and sex does not affect a person's perception of a climber's strength. Only the sex of the climber is significant. 

$H_A:$ The interaction between type of climbing and sex does affect a person's perception of a climber's strength. Only looking at sex is not a significant indicator of a climber's perceived strength and ability.

```{r}
#ANOVA
ANOVA_test <- lm(Rating ~ Type * Sex, data = simple_data)
anova(ANOVA_test)

pander(anova(ANOVA_test))
```

After looking at ANOVA output, we see that sex alone (p = 0.9) is not a significant predictor of a climber's perceived strength. In fact, the interaction between type of climbing and sex (p = 0.01) is statistically significant. Therefore, we reject the null hypothesis and conclude that the interaction between type of climbing and sex is statistically significant predictor of a climber's perceived stength and ability. 

# Residual Analysis 

_Check conditions on the residuals._

```{r}
hist(ANOVA_test$residuals)
```
The histogram of the residuals appears to follow a slightly normal distribution and for the sake of this exercise we will assume the condition of normality is met. 

```{r}
qqnorm(ANOVA_test$residuals)
```
The normal Q-Q Plot appears to be relatively linear, so we will assume the condition of normality has been met. 


# Equal Variance

```{r, message = F}
favstats(Rating ~ Sex|Type, data = simple_data)
```

The largest standard deviation (SD = 0.94) is not three times larger than the smallest standard deviation (SD = 0.71). Therefore Fisher's S assumption is met. 

# Assessing Third Factor: Prior Knowledge of Climbing

$H_0:$: The factor of prior knowledge of the sport of climbing will not have a statistically significant effect on the perceived strength of climbers of different sexes and on different types of climbs.  

$H_A:$ The factor of prior knowledge of the sport of climbing will have a statistically significant effect on the perceived strength of climbers of different sexes and on different types of climbs.  

```{r}
# second ANOVA 
mod2 <- lm(Rating ~ Type * Sex * Climber, data = simple_data)
anova(mod2)

pander(anova(mod2))
```

H_0 remains

```{r}
hist(mod2$residuals)
qqnorm(mod2$residuals)
```

Plots look decent

*How to calculate favstats on this model? Use manual SD calculations instead

```{r}
simple_data %>% 
  group_by(Type, Sex, Climber) %>% 
  summarise(
    sd = sd(Rating)
  )

```


