---
title: "Data Cleaning and Analysis"
author: "Hannah Snell and Vivienne Maxwell"
date: "2/9/2021"
output: 
  html_document:
    highlight: tango
    theme: cosmo
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: true
    df_print: kable
    
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 5, fig.height = 4, fig.align = 'center')

library(tidyverse)
library(mosaic)
library(gdata)

# import Qualtrics data 
mp2data <- read_csv("mp2data.csv")
```

# Data Cleaning and Analysis

```{r include=FALSE}
# data cleaning 
mp2data <- mp2data[-c(1, 2), ]
mp2data <- mp2data %>% 
  select(Finished, ResponseId, Q2, Q3, Q4, Q6, type, sex, climber) %>% 
  filter( 
    Finished != FALSE, 
    Q2 != "No",
    Q3 != "No, I Do Not Consent" 
    ) %>% 
  select(ResponseId, Q4, Q6, type, sex) %>% 
  mutate(type = tolower(type))

mp2data <- rename.vars(mp2data, from = c("Q4", "Q6", "type", "sex"),
         to = c("Climber", "Rating", "Type", "Sex"))
```

```{r echo=FALSE}
# starter plot
ggplot(mp2data, aes(x = Type, y = Rating, color = Sex)) +
  geom_point() + 
  geom_jitter(width = 0.2, height = 0.1) +
  facet_wrap(~ Climber) +
  labs(title = "Side-by-Side Dotplot: Can You Hang?", y  = "Overall Competency Rating of Climber") + 
  theme_bw()
```

```{r echo=FALSE}
# starter plot
ggplot(mp2data, aes(x = Type, y = as.numeric(Rating), color = Sex)) +
  geom_boxplot() + 
  facet_wrap(~ Climber) +
  labs(title = "Side-by-Side Boxplot: Can You Hang?", y  = "Overall Competency Rating of Climber") + 
  theme_bw()
```

```{r}
#anova data set
simple_data <- mp2data[-c(1)]
```

```{r}
write.csv(simple_data, "simple_data.csv")
```

**ANOVA Hypothesis** 

$H_0:$ Men are perceived as stronger, regardless of the type of climbing. Therefore, the interaction between type of climbing and sex does not affect a person's perception of a climber's strength. Only the sex of the climber is significant. 

$H_A:$ The interaction between type of climbing and sex does affect a person's perception of a climber's strength. Only looking at sex is not a significant indicator of a climber's perceived strength and ability.

```{r}
#ANOVA
ANOVA_test <- lm(Rating ~ Type * Sex, data = simple_data)

anova(ANOVA_test)
```

After looking at ANOVA output, we see that sex alone (p = 0.9) is not a significant predictor of a climber's perceived strength. In fact, the interaction between type of climbing and sex (p = 0.01) is statistically significant. Therefore, we reject the null hypothesis and conclude that the interaction between type of climbing and sex is statistically significant predictor of a climber's perceived stength and ability. 

#Residual Analysis 

Check conditions on the residuals

```{r}
hist(ANOVA_test$residuals)
```
The histogram of the residuals appears to follow a slightly normal distribution and for the sake of this exercise we will assume the condition of normality is met. 

```{r}
qqnorm(ANOVA_test$residuals)
```
The normal Q-Q Plot appears to be relatively linear, so we will assume the condition of normality has been met. 


#Equal Variance

```{r}
favstats(Rating ~ Sex|Type, data = simple_data)
```

The largest standard deviation (SD=0.94) is not three times larger than the smallest standard deviation (SD=0.71). Therefore Fisher's S assumption is met. 